Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering

### 2014-09-13 monitoring.tex
Various edits based on a full pass.

In this framework, a user specifies how much monitoring should be done in a form of
a target 
overhead budget, a target coverage, a percentage of monitoring operations to be performed, etc.
Then, the framework dynamically drops a portion of monitoring operations to match
the target. In particular, this paper focuses on matching a performance overhead target
while maximizing the coverage. Given that the overhead of custom hardware monitors can
already be quite low, the focus is on enabling the trade-off in {\em general-purpose} 
monitoring systems that support a wide range of monitors.
We also consider the target overhead as a soft constraint and do not aim to
provide a strict worst-case guarantee.

@@

{\bf Cooperative testing, debugging, and protection:}
Recently studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
or attack detection may be done in a cooperative fashion across
a large number of systems. In this case, each system is only willing to pay very low
overhead, say a few percents, and only performs a small subset of checks. 
A high coverage is achieved by having different systems check different parts of a program.
The partial monitoring framework allows high-overhead monitoring to be used in a
cooperative fashion. 
The main metric that represents the effectiveness of monitoring in this case is
the coverage (the percentage of checks there were performed) over multiple runs.

@@

{\bf Partial protection for low overhead:}
Even without real-time constraints, systems may have tight budgets for performance,
energy, or hardware overhead that they can tolerate for run-time monitoring. In such cases,
full monitoring cannot be enabled unless its overhead is low enough. The adjustable
monitoring allows partial protection on such systems. For example, array bounds
may be checked for a subset of memory accesses. For DIFT, a subset of information flows
may be tracked for partial attack detection. In this scenario, the effectiveness of 
partial monitoring can be measured as the percentage of run-time checks that are
performed on each program run, which reflects how likely for a bug or an attack 
to be detected for a system.

>>>>

In this framework, a user specifies how much monitoring should be done in the form of
a target 
overhead budget, a target coverage, a percentage of monitoring operations to be performed, etc.
Then, the framework dynamically drops a portion of monitoring operations to match
the target. In particular, this paper focuses on matching a performance overhead target
while maximizing the coverage. Given that the overhead of custom hardware monitors can
already be quite low, the focus is on enabling the trade-off in {\em general-purpose} 
monitoring systems that support a wide range of monitors.
We also consider the target overhead as a soft constraint and do not aim to
provide a strict worst-case guarantee.

@@

{\bf Cooperative testing, debugging, and protection:}
Recently, studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
or attack detection may be done in a cooperative fashion across
a large number of systems. In this case, each system is only willing to pay very low
overhead (e.g., a few percent) and only performs a small subset of checks. 
High coverage is achieved by having different systems check different parts of a program.
The partial monitoring framework allows high-overhead monitoring to be used in a
cooperative fashion. 
The main metric that represents the effectiveness of monitoring in this case is
the coverage (the percentage of checks there were performed) over multiple runs.

@@

{\bf Partial protection for low overhead:}
Even without real-time constraints, systems may have tight budgets for the performance,
energy, or hardware overhead that they can tolerate for run-time monitoring. In such cases,
full monitoring cannot be enabled unless its overhead is low enough. Adjustable
monitoring allows partial protection on such systems. For example, array bounds
may be checked for a subset of memory accesses. For DIFT, a subset of information flows
may be tracked for partial attack detection. In this scenario, the effectiveness of 
partial monitoring can be measured as the percentage of run-time checks that are
performed on each program run, which reflects how likely it is for a bug or an attack 
to be detected for a system.

### 2014-09-13 policies.tex
Various edits based on a full pass.

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components:
\begin{enumerate}
  \item When do we need to drop events in order to enable reduced overheads?
  \item Which events should be dropped?
\end{enumerate}

Section~\ref{sec:policies.when} discusses options for deciding when events are
dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
events can be dropped.

@@

Alternatively, we can specify a target overhead and estimate, at run-time, the
overheads of monitoring in order to decide whether dropping is needed.
% In order to decide whether a monitoring event should be dropped or not, we need
% a way to keep track of overheads. 
We present a specific method
for measuring execution time overheads. However, we note that other methods
could also be used in order to target budgets on IPC, power, energy, etc.  
% For example, one way to measure execution time overheads would be to insert
% checkpoints in the program. By recording the execution time of these
% checkpoints without monitoring, the overheads incurred while running with
% monitoring could be determined. Although this would give an accurate
% measurement of overheads, it would require modifying the program binary. In
% addition, the overhead budget for monitoring would only be updated at
% checkpoints.

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

@@

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
is decremented. It is difficult to perfectly determine when the main core is impacted by monitoring due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

@@

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dependence graph for metadata. Square nodes represent
    events where checks are performed.} 
    \label{fig:policies.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.
The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.
That is, by the time we decide to drop event {\tt E}, we have already 
updated metadata due to events {\tt C} and {\tt D} which is no longer needed.

\begin{figure}
  \begin{center}
  \subfloat[unrestricted dropping]{
    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
    \label{fig:policies.all_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[source-only dropping]{
    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
    \label{fig:policies.source_drop}
  }
  \vspace{-0.1in}
%   \\
%   \subfloat[sub-flow dropping]{
%     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
%     \label{fig:policies.min_drop}
%   }
  \end{center}
  \vspace{-0.1in}
  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
  dropping decisions are made.}
  \label{fig:policies.policies}
\end{figure}

@@

For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overheads, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.

>>>>

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components:
\begin{enumerate}
  \item When do we need to drop events in order to enable reduced overheads? (Section~\ref{sec:policies.when})
  \item Which events should be dropped? (Section~\ref{sec:policies.which})
\end{enumerate}

@@

Alternatively, we can specify a target overhead and estimate, at run-time, the
overheads of monitoring in order to decide whether dropping is needed.
The overhead budget is specified as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

@@

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on every $k$-th
cycle of the main core (e.g., every 5th cycle for a 20\% target budget). 
% This $k$ can be calculated by taking the reciprocal of
% the target budget. For example, if the target budget is 20\%, then the counter
% increments on every 5th clock cycle. 
The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, the
measured slack is decremented. It is difficult to precisely determine the
entire impact of monitoring on the main core due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

@@

% % Varying slack's impact on coverage for UMC
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
%     \vspace{-0.3in}
%     \caption{Example dependence graph for metadata. Square nodes represent
%     events where checks are performed.} 
%     \label{fig:policies.dataflow_graph}
%     \vspace{-0.1in}
%   \end{center}
% \end{figure}

\begin{figure}
  \begin{center}
  \subfloat[unrestricted dropping]{
    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
    \label{fig:policies.all_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[source-only dropping]{
    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
    \label{fig:policies.source_drop}
  }
  \vspace{-0.1in}
%   \\
%   \subfloat[sub-flow dropping]{
%     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
%     \label{fig:policies.min_drop}
%   }
  \end{center}
  \vspace{-0.1in}
  \caption{Comparison of dropping policies using metadata dependence graphs.
  Square nodes represent events where check are performed. Blue (dark) nodes
  indicate which nodes can be dropped.}
  \label{fig:policies.policies}
\end{figure}

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.
The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.
That is, by the time we decide to drop event {\tt E}, we have already 
updated metadata due to events {\tt C} and {\tt D} which is no longer needed.

@@

Depending on the target application of partial monitoring, different policies
are more applicable.
For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overheads, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.

### 2014-09-13 related.tex
Various edits based on a full pass.

% Monitoring
Many monitoring schemes have been developed for various security, reliability,
debugging, and other capabilities. There exist monitoring schemes for uninitialized
memory check \cite{mondrian-asplos02}, array bounds check
\cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
memory-related bugs and errors.  Control flow verification techniques perform
monitoring to check that program execution follows expected control flows
\cite{schuette-comp87, impres-dac06,
kayaalp-isca12}.  Our monitoring architecture is designed to
be generally applicable to any of these instruction-grained monitoring schemes.
% In particular, our proposed architecture may be especially useful for some of
% these more complex schemes which also have higher overheads.  Similarly,
% although many of these schemes showed low overheads in custom hardware, our
% architecture can be used to reduce the overheads when applying these schemes on
% a multi-core platform.

@@

% Sampling
Statistical sampling is a method that has been applied to various debugging
techniques in order to reduce their performance overheads. For example, Liblit
et al. sampled program runs of thousands of users in order to find bugs
\cite{liblit-pldi05}. This idea has also been applied to detecting memory
leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and dynamic information flow tracking\cite{testudo-micro08} and
other run-time monitoring techniques. 
These sampling-based techniques all limit the debugging capabilities in order
to provide lower performance overheads, similar to our architecture's goals.
% Limited Monitoring
In terms of partial monitoring with adjustable overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. 
Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring specifically for hard real-time systems.

Our work presents a general and efficient hardware platform to enable
partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
techniques as well as monitor implementations. By studying various potential
uses of partial monitoring, including sampling for cooperative debugging and
partial protection for low overheads, we have determined important dropping
policy decisions to consider in order to best meet the requirements of the
application.

>>>>

% Monitoring
% Many monitoring schemes have been developed for various security, reliability,
% debugging, and other capabilities. There exist monitoring schemes for uninitialized
% memory check \cite{mondrian-asplos02}, array bounds check
% \cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
% tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
% Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
% memory-related bugs and errors.  Control flow verification techniques perform
% monitoring to check that program execution follows expected control flows
% \cite{schuette-comp87, impres-dac06,
% kayaalp-isca12}.  Our monitoring architecture is designed to
% be generally applicable to any of these instruction-grained monitoring schemes.

% In particular, our proposed architecture may be especially useful for some of
% these more complex schemes which also have higher overheads.  Similarly,
% although many of these schemes showed low overheads in custom hardware, our
% architecture can be used to reduce the overheads when applying these schemes on
% a multi-core platform.

@@

% Sampling
Statistical sampling is a method that has been applied to various debugging
techniques in order to reduce their performance overheads. For example, Liblit
et al. sampled program runs of thousands of users in order to find bugs
\cite{liblit-pldi05}. This idea has also been applied to detecting memory
leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and dynamic information flow tracking\cite{testudo-micro08} and
other run-time monitoring techniques. 
These sampling-based techniques all limit the debugging capabilities in order
to provide lower performance overheads, which is one of the target applications of our architecture.
% Limited Monitoring
In terms of partial monitoring with adjustable overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. 
Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring specifically for hard real-time systems.

In contrast, our work is the first that we know of to present a general hardware platform that enables
partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
techniques as well as monitor implementations (core, FPGA, etc.). By studying various potential
uses of partial monitoring, including sampling for cooperative debugging and
partial protection for low overheads, we have determined important dropping
policy decisions to consider in order to best meet the requirements of the
application.

### 2014-09-13 tabs/previous_overheads.tex
Various edits based on a full pass.

DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline

>>>>

% DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
% FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
% Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
% Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline

DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.01x / 1.23x \\ \hline
FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.01x-1.04x / 1.09x \\ \hline
Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 1.05x-1.09x / 1.22x \\ \hline
Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1.01x-1.10x / 1.20x \\ \hline\hline

### 2014-09-13 title.tex
Various edits based on a full pass.

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of computer
systems. However, these run-time monitors can introduce large performance and energy
overheads, especially when performed on programmable cores.
%Traditionally, these overheads have been an
%all-or-nothing cost; monitoring cannot be used if the overheads are considered
%too large.  
In this paper, we introduce a hardware dataflow tracking engine that enables
adjustable overheads through partial monitoring. This allows a trade-off to be
made between monitoring coverage and overhead. This dataflow engine
can also be extended to filter out monitoring operations associated with null
metadata in order to reduce overheads.
% In this paper, we introduce a hardware dataflow tracking engine that can be
% used to filter out unnecessary monitoring and enable adjustable partial monitoring.
% The dataflow engine can identify events with null metadata so that they can
% be filtered out. To further reduce the overhead of monitoring, we propose
% to enable a trade-off between monitoring coverage and overhead by dropping certain
% monitoring operations. For this partial monitoring, the dataflow engine is used
% to track dropped monitoring flows so that false positives can be avoided.
Given this architecture, we investigate how the dropping decisions should be
made for partial monitoring and show that there exist interesting policy
decisions which create a trade-off between closely
matching an overhead target and the coverage that can be achieved.
Our experimental results show that overheads can be reduced significantly 
by trading off coverage. For example, although average overheads for full
monitoring range between 3-4x, the proposed architecture can still
achieve 43-82\% coverage when the slowdown is reduced to 1.5x.

>>>>

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of computer
systems. However, these run-time monitors can introduce large performance and energy
overheads, especially for flexible systems that support a range of monitors.
%Traditionally, these overheads have been an
%all-or-nothing cost; monitoring cannot be used if the overheads are considered
%too large.  
In this paper, we introduce a hardware dataflow tracking engine that enables
adjustable overheads through partial monitoring. This allows a trade-off to be
made between monitoring coverage and overhead. This dataflow engine
can also be extended to filter out monitoring operations associated with null
metadata in order to reduce overheads.
% In this paper, we introduce a hardware dataflow tracking engine that can be
% used to filter out unnecessary monitoring and enable adjustable partial monitoring.
% The dataflow engine can identify events with null metadata so that they can
% be filtered out. To further reduce the overhead of monitoring, we propose
% to enable a trade-off between monitoring coverage and overhead by dropping certain
% monitoring operations. For this partial monitoring, the dataflow engine is used
% to track dropped monitoring flows so that false positives can be avoided.
Given this architecture, we investigate how the dropping decisions should be
made for partial monitoring and show that there exist interesting policy
decisions depending on the target application of partial monitoring.
Our experimental results show that overheads can be reduced significantly 
by trading off coverage. For example, with average overheads of 3-4x for full
monitoring range, the proposed architecture can still
achieve 43-82\% coverage when the slowdown is reduced to 1.5x.
